---
title: "Kontur Population"
author: "Orhan Aktas"
date: '2022-11-24'
output:   
    html_document:
      df_print: !expr knitr::kable
      toc: yes
      number_sections: yes
      toc_depth: 2
---

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  class.source = "bg-warning text-warning"
)
```

#### Required Libraries {.unlisted .unnumbered}

```{r libraries, message=FALSE, warning=FALSE}
library(osrm)
library(sf)
library(tidyverse)
library(leaflet)
```

# Introduction

It would be nice to extract only the hexagons located within the boundaries of Istanbul.


# Import data

## population data for turkey
This data was previously downloaded from https://data.humdata.org/dataset/kontur-population-turkiye.I just read this file initially. After selecting the hexagons for Istanbul, I don't read it.
```{r eval=FALSE}

nc = read_sf("data/kontur_pop_tr.gpkg")

```

## boundries for turkey

This data was previously downloaded from https://data.humdata.org/dataset/kontur-boundaries-turkey.
```{r}

boundries_tr <- read_sf("data/kontur_boundaries_TR_20220407.gpkg")


```

map
```{r}
boundries_tr |> filter(admin_level==4)|> st_geometry() |> plot()
boundries_tr |> filter(admin_level==4)|> select(population) |> plot()
```


## boundry for istanbul
```{r}
boundry_ist<-boundries_tr |> filter(name_en=="Istanbul")

boundries_tr |> 
  filter(admin_level==10) |> 
  st_join(boundry_ist) |> 
  filter(name_en.y=="Istanbul") |> 
  select(population.x)|> 
  plot()

```










prepare hexagons for Istanbul and write to a separate file. This was done initially.
```{r eval=FALSE}
crs_string <- '+proj=longlat +datum=WGS84'

hex_ist <- nc |> 
  st_transform(4326) |> 
  st_intersection(boundry_ist  |> st_transform(4326)) |> 
  st_transform(crs = crs_string)

write_rds(hex_ist.rds,"data/hex_ist.rds")
```

read previously prepared file for Istanbul
```{r}
hex_ist <- read_rds("data/hex_ist.rds")
```


map with ggplot
```{r}
ggplot() +
  geom_sf(data = hex_ist , aes(fill = population),size=0.001) +
  theme(
    legend.position = "none"
  )+
  scale_color_distiller(palette = "Spectral") +
  labs(
    caption = "Population Density for 400m H3 Hexagons\nSource: Kontur Population\nGraphic: Orhan Aktaş"
  ) +
  theme_void() +
  theme(
    legend.position = "none"
  )
```

leaflet example for examining grids
```{r}

# Create a continuous palette function
pal <- colorNumeric(
  palette = "Blues",
  domain = hex_ist$population)

# Apply the function to provide RGB colors to addPolygons
leaflet(hex_ist) %>% 
     addProviderTiles("CartoDB.PositronNoLabels") %>% 
     addPolygons(stroke = TRUE, weight=0.4,smoothFactor = 0.2, fillOpacity = 0,
    color ="black" )
```

leaflet example for examining populations
```{r}

# Create a continuous palette function
pal <- colorNumeric(
  palette = "Blues",
  domain = hex_ist$population)

# Apply the function to provide RGB colors to addPolygons
leaflet(hex_ist) %>% 
     addProviderTiles("CartoDB.PositronNoLabels") %>% 
     addPolygons(stroke = TRUE, weight=0.1,smoothFactor = 0.2, fillOpacity = 0.5,
    color ="black" ,fillColor =~pal(population) )
```