---
title: "Kontur Population"
author: "Orhan Aktas"
date: '2022-11-24'
output:   
    html_document:
      df_print: !expr knitr::kable
      toc: yes
      number_sections: yes
      toc_depth: 2
---

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  class.source = "bg-warning text-warning"
)
```

#### Required Libraries {.unlisted .unnumbered}

```{r libraries, message=FALSE, warning=FALSE}
library(osrm)
library(sf)
library(tidyverse)
library(leaflet)
```

# Introduction

It would be nice to extract only the hexagons located within the boundaries of Istanbul.


# Import data

## population data for turkey
This data was previously downloaded from https://data.humdata.org/dataset/kontur-population-turkiye.I just read this file initially. After selecting the hexagons for Istanbul, I don't read it.
```{r eval=FALSE}

nc = read_sf("data/kontur_pop_tr.gpkg")

```

## gis

hgm
```{r}
hgm_boundary_turkey<-  read_sf("data/gis/hgm_boundary/Ülke_Sınırı.shp")
hgm_boundary_city<-  read_sf("data/gis/hgm_boundary/İl_Sınırı.shp")
hgm_boundary_district<-  read_sf("data/gis/hgm_boundary/İlçe_Sınırı.shp")
hgm_boundary_settlement<-  read_sf("data/gis/hgm_boundary/Yerleşim_Nokta.shp")



```


districts
```{r}
districts <-  read_sf("data/gis/districts/Export_Output_2.shp")
districts |> st_drop_geometry()|> select(name=ILCE_ADI,population=NUFUS) |> head()
districts |> st_drop_geometry()|>summarise(total_pop=sum(NUFUS))
districts |> select(NUFUS) |> plot()
```

neighbourhoods
```{r}
neighbourhoods <-  read_sf("data/gis/neighbourhoods/Export_Output.shp")
neighbourhoods |> st_drop_geometry()|> select(name=MAHALLE__1,population=NUFUS)|> head()
neighbourhoods |> st_drop_geometry()|>summarise(total_pop=sum(NUFUS))
neighbourhoods |> select(NUFUS) |> plot()
```



Bunu tam anlayamadım.
```{r}

nufus <-  read_sf("data/gis/nufus/Export_Output.shp")

```

what is oran?
It might be a percentage between different areas. There are two different area variables. Lets try to calculate same values again and map the old and the new column. If they are the same, then we can understand its meaning.
```{r}
nufus |> 
  select(AREA,Shape_Ar_1,oran) |> 
  st_drop_geometry() |> 
  mutate(new_oran=Shape_Ar_1*10^12/AREA) |> 
  ggplot(aes(x=oran,y=new_oran))+
    geom_point()
```



stations of railway and metrobus
```{r}
stations <-  read_sf("data/gis/isbak/ttb.shp")
stations |> head()
stations |> names()
stations |> st_geometry() |> plot()
```



catchment areas by distance
```{r}
ttb2 <-  read_sf("data/gis/ttb/ttb_3duzey.shp")
ttb2 |> head()
ttb2 |> select(FromBreak) |> plot()
```




map
```{r}
nufus |>   select(Sum_Outp_3)   |> plot()
# nufus |>  st_geometry() |> plot()
#boundaries_tr |> filter(admin_level==4)|> select(population) |> plot()
```

## boundaries for turkey

This data includes aggregated population values by administrative divisions. It was previously downloaded from https://data.humdata.org/dataset/kontur-boundaries-turkey.

import data
```{r}
boundaries_tr <- read_sf("data/kontur_boundaries_TR_20220407.gpkg")
```

map
```{r}
boundaries_tr |> filter(admin_level==4)|> st_geometry() |> plot()
boundaries_tr |> filter(admin_level==4)|> select(population) |> plot()
```


## boundary for istanbul

create 
```{r}
boundary_ist<-boundaries_tr |> filter(name_en=="Istanbul")
```

create neighbourhood boundaries for Istanbul
```{r}
boundaries_ist<-boundaries_tr |> 
  filter(admin_level==10) |> 
  st_join(boundary_ist) |> 
  filter(name_en.y=="Istanbul")
  
boundaries_ist |> 
  select(population.x)|> 
  plot()
```










prepare hexagons for Istanbul and write to a separate file. This was done initially.
```{r eval=FALSE}
crs_string <- '+proj=longlat +datum=WGS84'

hex_ist <- nc |> 
  st_transform(4326) |> 
  st_intersection(boundary_ist  |> st_transform(4326)) |> 
  st_transform(crs = crs_string)

write_rds(hex_ist.rds,"data/hex_ist.rds")
```

read previously prepared file for Istanbul
```{r}
hex_ist <- read_rds("data/hex_ist.rds")
```


map with ggplot
```{r}
ggplot() +
  geom_sf(data = hex_ist , aes(fill = population),size=0.001) +
  theme(
    legend.position = "none"
  )+
  scale_color_distiller(palette = "Spectral") +
  labs(
    caption = "Population Density for 400m H3 Hexagons\nSource: Kontur Population\nGraphic: Orhan Aktaş"
  ) +
  theme_void() +
  theme(
    legend.position = "none"
  )
```

leaflet example for examining grids
```{r}

# Create a continuous palette function
pal <- colorNumeric(
  palette = "Blues",
  domain = hex_ist$population)

# Apply the function to provide RGB colors to addPolygons
leaflet(hex_ist) %>% 
     addProviderTiles("CartoDB.PositronNoLabels") %>% 
     addPolygons(stroke = TRUE, weight=0.4,smoothFactor = 0.2, fillOpacity = 0,
    color ="black" )
```

leaflet example for examining populations
```{r}

# Create a continuous palette function
pal <- colorNumeric(
  palette = "Blues",
  domain = hex_ist$population)

# Apply the function to provide RGB colors to addPolygons
leaflet(hex_ist) %>% 
     addProviderTiles("CartoDB.PositronNoLabels") %>% 
     addPolygons(stroke = TRUE, weight=0.1,smoothFactor = 0.2, fillOpacity = 0.5,
    color ="black" ,fillColor =~pal(population) )
```


## Comparison with data from IMM (Istanbul Metropolitan Municipality)

join two neighbourhood data
```{r}
comparison <-neighbourhoods  |> 
  st_transform(4326) |> 
  select(MAHALLE__1,pop_real=NUFUS) |> 
  st_join(
    boundaries_ist |> 
      select(pop_kntr=population.x,mah_kntr=name.x) |> 
      mutate(mah_kntr = toupper(str_replace(mah_kntr, "Mahallesi", ""))) |> 
      st_centroid()
  ) 
```

take a glance to the table
```{r}
comparison |> head(20)
```

plot neighbourhoods based on two populations separately.
```{r}
comparison |> select(pop_real,pop_kntr) |> plot()
```

lastly, create a scatter chart to compare two population data
```{r}
comparison |> 
  st_drop_geometry() |>
  mutate(
    pct_real=pop_real/sum(pop_real)*1000,
    pct_kntr=pop_kntr/sum(pop_kntr,na.rm = TRUE)*1000
  ) |> 
  ggplot(aes(x=pct_real,y=pct_kntr))+
  geom_point()
```

I'm afraid these two data sources are not similar. since one was obtained from the municipality, I am concluding that the hexagon data does unfortunately not reflect the real situation.
